---
title: "Visualisation in R Part 3 - Lab 8"
output: html_notebook
---



#  Visualisation in R Part 3
## Lab 8

```{r}
warning=FALSE
```



```{r}
# Load libraries
library(readr)
library(ggplot2)
library(sqldf)
library(viridis)
library(hrbrthemes)
library(dplyr)
library(tidyr)
```


### Initial Data Load/Cleaning Routines

```{r}
# Read CSV file with Olympoic Data
olympicdata<- read.csv( 'OlympicGames.csv' , sep= ',' , header=T)


# Build dataset with gold medals per country
resultsmedalsgold<-sqldf("select country, 
                          count(Medal) as gold 
                          from olympicdata where 
                          Medal=='gold' 
                          group by country")

# Build dataset with silver medals per country
resultsmedalssilver<-sqldf("select country, 
                            count(Medal) as silver 
                            from olympicdata 
                            where Medal=='silver' 
                            group by country") 

# Build dataset with bronze medals per country
resultsmedalsbronze<-sqldf("select country, 
                            count(Medal) as bronze 
                            from olympicdata 
                            where Medal=='bronze' 
                            group by country")

# Combine medals dataframes
resultsmedals<-merge(resultsmedalsgold,resultsmedalssilver,by="Country",all=TRUE) 
resultsmedals2<-merge(resultsmedals,resultsmedalsbronze,by="Country",all=TRUE)
```


### Visualization One - What country has won most Silver medals since 2000? (2 marks)


```{r}
# SQL dataframe generation
dfMostSilverMedalsCountry<-sqldf("select country, 
                            count(Medal) as Silver 
                            from olympicdata 
                            where Medal=='silver' 
                            and Year >= 2000
                            group by country
                            order by Silver desc
                            limit 7") 

# Display dataframe
dfMostSilverMedalsCountry

# Visualization 1 - Horizontal Bar Chart of top performing silver medalist countries
# since 2000 (Top 7)
gg1 <- ggplot(data=dfMostSilverMedalsCountry, aes(x=reorder(Country, +Silver), y=Silver)) +
  geom_bar(stat="identity", colour="black", fill="blue") +
  labs(y = "Count of Silver Medals", x = "Country Name") + 
  ggtitle('Countries With Most Silver Medals Since 2000 (Top 7)') +
  theme(plot.title = element_text(size = 17),
        axis.title = element_text(size = 15),
        axis.text.y = element_text(size=12, face="bold", colour = "black")) +
  coord_flip() + 
  geom_label(
    aes(label = round(Silver)), 
    #hjust=1.15,
    size = 2.5, fontface = "bold",
    ## turn into white box without outline
    fill = "white", label.size = 0.5)
print(gg1)
```

# Visualization Two - How is the gender balance among the United States gold medalist? (2 marks)

```{r}
# SQL dataframe generation
dfUSGoldMedalsGenderBal<-sqldf("select gender, 
                            count(Medal) as Gold 
                            from olympicdata 
                            where Medal=='gold' 
                            and country=='USA'
                            group by gender
                            order by Gold desc") 

# Display dataframe
dfUSGoldMedalsGenderBal


# Visualization Two  Bar Graphs - Us Gold Medal Gender Balance
gg2 <- ggplot(dfUSGoldMedalsGenderBal,aes(x=reorder(Gender, -Gold), y = Gold, fill = Gender)) + 
  geom_bar(position = "stack", stat = "identity", colour="black") + 
  ggtitle("Gender Mix - US Gold Medalists") + 
  labs(y = "Gold Medals", x="Gender") + 
  theme(legend.position = "none",
        plot.title = element_text(size = 17),
        axis.title = element_text(size = 15),
        axis.text.y = element_text(size=10, face="bold", colour = "black"),
        axis.text.x = element_text(size=10, face="bold", colour = "black")) +
  geom_label(
    aes(label = Gold), 
    size = 5, fontface = "bold",
    fill = "white", label.size = 1)  
print(gg2)
```

# Visualization Three - What are the best sports for Sweden, USA, Austria and Switzerland? (2 marks)

```{r}
# SQL dataframe generation - four seperate SQL queries for each country
dfBestSportsSweden<-sqldf("select Country, Sport, 
                            count(Medal) as Medal_Count 
                            from olympicdata 
                            where country =='Sweden'
                            group by Sport
                            order by Medal_Count desc
                            limit 3") 


dfBestSportsUSA<-sqldf("select Country, Sport, 
                            count(Medal) as Medal_Count 
                            from olympicdata 
                            where country =='USA'
                            group by Sport
                            order by Medal_Count desc
                            limit 3") 


dfBestSportsSwitzerland<-sqldf("select Country, Sport, 
                            count(Medal) as Medal_Count 
                            from olympicdata 
                            where country =='Switzerland'
                            group by Sport
                            order by Medal_Count desc
                            limit 3") 


dfBestSportsAustria<-sqldf("select Country, Sport, 
                            count(Medal) as Medal_Count 
                            from olympicdata 
                            where country =='Austria'
                            group by Sport
                            order by Medal_Count desc
                            limit 3") 


# Merge dataframes 
resultsBestSports<-merge(dfBestSportsSweden,dfBestSportsUSA,all=TRUE) 
resultsBestSports<-merge(resultsBestSports,dfBestSportsSwitzerland,all=TRUE)
resultsBestSports<-merge(resultsBestSports,dfBestSportsAustria,all=TRUE)  



# Visualization Three - Facet Wrap Bar Charts 
gg3 <- ggplot(resultsBestSports,aes(x = reorder(Sport,-Medal_Count), y = Medal_Count, fill=Sport)) + 
  geom_bar(position = "stack", stat = "identity", colour="black") + 
  ggtitle("Best Sports - Austria, Sweden, Switzerland, USA") + 
  facet_wrap(~Country, nrow = 4) + 
  labs(y = "Medal Count For Each Country", x = "Sport") + 
  theme_ipsum()  +
  theme(legend.position = "top",
        plot.title = element_text(size = 15),
        axis.title = element_text(size = 10),
        axis.text.y = element_text(size=10, face="bold", colour = "black")) 
print(gg3)

```

# Visualization Four - Who has the most total medals? (2 marks)

```{r}

# Pivot table to prepare data for bar chart
resultsMedals_Reshape <- resultsmedals2 %>% 
  pivot_longer(c(gold, silver, bronze), # values to pivot or reshape
               names_to = "Medal_Type", # Rename column for exam type (written or oral)
               values_to = "Medal_Number") # Re-name column containing the exam scores



# SQL to build dataframes
dfTop5MedalWinners<-sqldf("select Country,
                            sum(Medal_Number) as Total_Medal_Count 
                            from resultsMedals_Reshape 
                            group by Country
                            order by Total_Medal_Count desc
                            limit 7") 


dfTopMedalWinners<-sqldf("select Country, Medal_Type,
                            sum(Medal_Number) as Total_Medal_Count 
                            from resultsMedals_Reshape
                            where country in 
                              (select distinct Country from dfTop5MedalWinners)
                            group by Country, Medal_Type
                            order by Country, Total_Medal_Count desc") 


#convert 'position' to factor and specify level order
dfTopMedalWinners$Medal_Type <- factor(dfTopMedalWinners$Medal_Type, levels=c('gold', 'silver', 'bronze'))



# Visualization Four -Stacked Bar Char of Medals
ggplot(dfTopMedalWinners,aes(x=reorder(Country,-Total_Medal_Count), fill=Medal_Type)) +
  geom_bar(aes(weight=Total_Medal_Count),position="stack") +
  scale_fill_manual(values=c('gold', 'light blue', 'tomato')) +
  geom_text(position="stack",
            aes(Country,
                Total_Medal_Count,
                label=Medal_Type),
            vjust = 1.5,
            size=5) + 
  ggtitle('Top 7 Countries by Olympic Medals Won') +
  labs(y = "Total Medal Count For Each Country", x = "Country") +
  theme(legend.position = "none",
        plot.title = element_text(size = 17),
        axis.title = element_text(size = 15),
        axis.text.y = element_text(size=12, face="bold", colour = "black"),
        axis.text.x = element_text(size=12, face="bold", colour = "black"))
```
# Visualization Five - What is the variation and spread of ages amongst gold and silver medalists? (2 marks)

```{r}
# Remove Rows with missing values for Age
dfOlympicAgesCleaned<-na.omit(olympicdata)

# Build datasets with Gold and Silver Medal Winners by Age
dfAgesGoldWinners<-sqldf("select Athlete_Age, 
                            'Gold' as Medal_Rank
                            from dfOlympicAgesCleaned 
                            where Medal=='gold' 
                            group by Athlete_Age
                            order by Athlete_Age desc") 


dfAgesSilverWinners<-sqldf("select Athlete_Age, 
                            'Silver' as Medal_Rank
                            from dfOlympicAgesCleaned 
                            where Medal=='silver' 
                            group by Athlete_Age
                            order by Athlete_Age desc") 


# Combine dataframes of Gold and Silver Medal winners by age
dfAgesGold_and_SilverWinners<-merge(dfAgesSilverWinners,dfAgesGoldWinners,all=TRUE) 


# Convert the variable Medal Type from a numeric to a factor variable
dfAgesGold_and_SilverWinners$Medal_Rank <- as.factor(dfAgesGold_and_SilverWinners$Medal_Rank)


# Visualization Five -  Box Plot of Silver Medalists
ggplot(dfAgesGold_and_SilverWinners, aes(x=Medal_Rank, y=Athlete_Age, fill=Medal_Rank)) + 
  geom_boxplot() +
  labs(y = "Age", 
       x = "Athlete Medal Winners - Age Distribution", 
       fill='Medal Ranks') + 
  ggtitle('BoxPlot - Distribution of Ages - Gold + Silver Medalists') +
  theme(legend.position = "right",
        plot.title = element_text(size = 17),
        axis.title = element_text(size = 15),
        axis.text.x = element_text(size=15, face="bold", colour = "black"))
```

