---
title: "R Notebook"
output: html_notebook
---


# Lab - Week 8

The visualisations are placed underneath each question.

```{r}

# Set Up Libraries
library(ggplot2)
library(viridis)
library(hrbrthemes)
library(sqldf)
library(tidyr) # tidy data

```


### 1. What are the total marks (oral plus written divided by two) for each student for each subject? (2 marks)

```{r}

# Impute missing values
avgmark <- sqldf("select AVG(Mark_Oral) from studentresult where Name = 'Mary Healy' AND Mark_Oral is not 'NA'")
avgmark

studentresult$Mark_Oral <- ifelse(is.na(studentresult$Mark_Oral), as.numeric(avgmark), studentresult$Mark_Oral)


# SQL to Set Up Exercise 1 + Exercise 2

# Convert date to a "real date" - Exercise 2
studentresult$DOB <- as.Date(studentresult$DOB, '%d-%m-%Y')

# Add today's date, also as a "real date" - Exercise 2
studentresult$today <- as.Date(Sys.Date())


# Average Result - Exercise 1
avgTotalResults <- sqldf('select Name, subject, 
                                      sum(Mark_Written) as Sum_Written_Mrks, 
                                      sum(Mark_Oral) as Sum_Oral_Mks,
                                      count(year) as Num_Years,
                                      (((sum(Mark_Written)+sum(Mark_Oral))/2)/count(year)) as Overall_Avg,
                                      floor((today-DOB)/365.25) as Age
                                      from studentresult group by Name, subject')

View(avgTotalResults)

# Question 1 Bar Graphs

gg <- ggplot(avgTotalResults,aes(x = Name, y = Overall_Avg, fill = Subject, 
                                 label = round(Overall_Avg, digits = 0)))
gg <- gg + geom_bar(position = "stack", stat = "identity", width=0.35)
gg <- gg + scale_fill_viridis(discrete = T, option = "E")
gg <- gg + ggtitle("Student Body Marks") 
gg <- gg + facet_wrap(~Subject, nrow = 3, scales = "fixed")
gg <- gg + theme_ipsum()
gg <- gg + labs(y = "Average Marks For Each Student")
gg <- gg + geom_label(
  aes(label = round(Overall_Avg, digits = 0)), 
  vjust=1.15,
  size = 3, fontface = "bold", family = "Fira Sans",
  ## turn into white box without outline
  fill = "white", label.size = 0)
gg <- gg + theme(legend.position = "none")
print(gg)


```


## 2. What is the relationship between age and mark? (2 marks)

```{r}

# Perform calculation for age
studentMarkProfile <- sqldf("select DISTINCT(Name), subject, Age, 
                             Overall_Avg as Average_Grade from avgTotalResults
                             order by subject")

View(studentMarkProfile)


# Question 2 Point Graphs
gg2 <- ggplot(data=studentMarkProfile, aes(x=Age, y= Average_Grade, group=1)) +
  geom_line() +
  geom_point(colour="black", size=2, shape=21, fill="white") +
  facet_wrap(~Subject, nrow = 1, scales = "free") +
  labs(y = "Average Marks For Each Student")
print(gg2)




```
### With the exception of Irish, mark scores drop as the student gets older.



## 3. Did any students do better on their written compared with their oral (or vice versa)? (2 marks)

```{r}

# Build SQL for Mary Healy
studentMarkComps <- sqldf("select Name, subject, Year,
                             Mark_Written,
                             Mark_Oral
                             from studentresult
                             where Name = 'Mary Healy'
                             order by year asc")


View(studentMarkComps)

studentMarkComps_Reshape <- studentMarkComps %>% 
  pivot_longer(c(Mark_Written, Mark_Oral), # values to pivot or reshape
               names_to = "Exam_Type", # names will be
               values_to = "Mark_Score") # values will be

studentMarkComps_Reshape # replay myReshapedData


# Question 3 Bar Graphs - Mary Healy
gg3a <- ggplot(data=studentMarkComps_Reshape, aes(x=Year, y=Mark_Score, fill=Exam_Type)) +
       geom_bar(stat="identity", position=position_dodge(), colour="black") +
       scale_fill_manual(values=c("#999999", "#E69F00")) +
       labs(y = "Student Marks") + 
       facet_wrap(~Subject, nrow = 1, scales = "fixed") +
       ggtitle('Mary Healy') 
print(gg3a)



# Build SQL for Hercule Poirot
studentMarkComps <- sqldf("select Name, subject, Year,
                             Mark_Written,
                             Mark_Oral
                             from studentresult
                             where Name = 'Hercule Poirot'
                             order by year asc")


View(studentMarkComps)

studentMarkComps_Reshape <- studentMarkComps %>% 
  pivot_longer(c(Mark_Written, Mark_Oral), # values to pivot or reshape
               names_to = "Exam_Type", # names will be
               values_to = "Mark_Score") # values will be

studentMarkComps_Reshape # replay myReshapedData


# Question 3 Bar Graphs - Hercule Poirot
gg3b <- ggplot(data=studentMarkComps_Reshape, aes(x=Year, y=Mark_Score, fill=Exam_Type)) +
  geom_bar(stat="identity", position=position_dodge(), colour="black") +
  scale_fill_manual(values=c("#009999", "#0000FF")) +
  labs(y = "Student Marks") + 
  facet_wrap(~Subject, nrow = 1, scales = "fixed") +
  ggtitle('Hercule Poirot') 
print(gg3b)




# Build SQL for Joe O'Neil
studentMarkComps <- sqldf("select Name, subject, Year,
                             Mark_Written,
                             Mark_Oral
                             from studentresult
                             where Name = 'Joe O''Neil'
                             order by year asc")


View(studentMarkComps)

studentMarkComps_Reshape <- studentMarkComps %>% 
  pivot_longer(c(Mark_Written, Mark_Oral), # values to pivot or reshape
               names_to = "Exam_Type", # names will be
               values_to = "Mark_Score") # values will be

studentMarkComps_Reshape # replay myReshapedData


# Question 3 Bar Graphs - Joe O'Neil
gg3c <- ggplot(data=studentMarkComps_Reshape, aes(x=Year, y=Mark_Score, fill=Exam_Type)) +
  geom_bar(stat="identity", position=position_dodge(), colour="black") +
  labs(y = "Student Marks") + 
  facet_wrap(~Subject, nrow = 1, scales = "fixed") +
  ggtitle("Joe O''Neil") 
print(gg3c)





```
### All students have done better in some years on Written compared to Oral


## 4. What subject obtained the best results on average? (2 marks)

```{r}

bestSubjectsAvgTbl <- sqldf("select sum(Mark_Written+Mark_Oral)/(count(Mark_Written)+count(Mark_Oral)) 
                                    as Subject_Average,  
                                    subject
                                    from studentresult
                                    group by subject")


View(bestSubjectsAvgTbl)


# Question 4 Bar Graphs - Best subject on Average
gg4 <- ggplot(data=bestSubjectsAvgTbl, aes(x=reorder(Subject, +Subject_Average), 
                                           y=Subject_Average, fill=Subject)) +
  geom_bar(stat="identity", position=position_dodge(), colour="black") +
  labs(y = "Average Subject Marks", x = "Subject") + 
  ggtitle('Best Performing Subject (on Average)') + 
  geom_label(
    aes(label = paste(round(Subject_Average, digits = 0),"%")), 
    vjust=1.15,
    size = 3, fontface = "bold",
    fill = "white", label.size = 0) +
  theme(legend.position = "none")
print(gg4)







```
### Irish achieved best results on average


## 5. What are the average results in oral exams across all subjects and all years per student? (2 marks)

```{r}

avgOralScore <- sqldf("select sum(Mark_Oral)/count(Mark_Oral)
                                    as Oral_Subject_Average,  
                                    Name
                                    from studentresult
                                    group by Name")


View(avgOralScore)


# Question 5 Bar Graphs - Best subject on Average
gg5 <- ggplot(data=avgOralScore, aes(x=reorder(Name, +Oral_Subject_Average), 
                                           y=Oral_Subject_Average, fill=Name)) +
  geom_bar(stat="identity", position=position_dodge(), colour="black") +
  labs(y = "Average Subject Marks in Oral Exams", x = "Student", face="bold") + 
  ggtitle('Student Performance in Oral Exams') + 
  geom_label(
    aes(label = paste(round(Oral_Subject_Average, digits = 0),"%")), 
    vjust=1.15,
    size = 3, fontface = "bold",
    fill = "white", label.size = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(size=12, face="bold", colour = "black"))
print(gg5)



```
# Avg Oral Scores; Mary = 81%, Joe = 70%, Hercule = 50%



