---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


```{r}
warning=FALSE
```
###
###
###
# Lab - Week 10

# Recreate given graphs using RStudio and ggplot


```{r}
# Load libraries
library(ggplot2)
library(viridis)
library(hrbrthemes)
library(sqldf)
library(tidyr) # tidy data


# Impute missing values in Student Record
avgmark <- sqldf("select AVG(Mark_Oral) from studentresult where Name = 'Mary Healy' AND Mark_Oral is not 'NA'")
avgmark

studentresult$Mark_Oral <- ifelse(is.na(studentresult$Mark_Oral), as.numeric(avgmark), studentresult$Mark_Oral)
```



###
###

#### Exercise 1: Looking at last week’s studentresults file, 
#### plot the results for the written and oral exams in two bars per student. (4 marks)

```{r}

###############################################################################################################
# Exercise 1: Looking at last week’s studentresults file, 
# plot the results for the written and oral exams in two bars per student. (4 marks)
###############################################################################################################


# Build SQL for all students average marks
avgOralScore <- sqldf("select Name,
                       sum(Mark_Written)/count(Mark_Written) as markw,
                       sum(Mark_Oral)/count(Mark_Oral) as marko
                       from studentresult
                       group by Name")

# Review aggregations of average score for students
View(avgOralScore)

# Pivot table to prepare data for bar chart
studentMarkComps_Reshape <- avgOralScore %>% 
  pivot_longer(c(markw, marko), # values to pivot or reshape
               names_to = "Exam_Type", # Rename column for exam type (written or oral)
               values_to = "Mark_Score") # Re-name column containing the exam scores


# Change data order to adjust graph legend
studentMarkComps_Reshape$Exam_Type <- factor(studentMarkComps_Reshape$Exam_Type, levels = c("markw","marko"))

# Display dataset
studentMarkComps_Reshape

# Question 1 Bar Graphs - Compare Average Student Results
# Order by ascending marks values
gg1 <- ggplot(data=studentMarkComps_Reshape, aes(x=reorder(Name, +Mark_Score), y=Mark_Score, fill=Exam_Type)) +
  geom_bar(stat="identity", position=position_dodge(), colour="black") +
  labs(y = "Student Marks (Average)", x = "Student Name") + 
  ggtitle('Average Student Performance Across Exam Types') +
  theme(legend.position = "right",
        plot.title = element_text(size = 17),
        axis.title = element_text(size = 15),
        axis.text.x = element_text(size=12, face="bold", colour = "black")) + 
  scale_fill_discrete(labels=c('Marks - Written', 'Marks - Oral'),name = "Exam Type") 
print(gg1)

```







#### Exercise 2: Plot one variable counting the instances that fall in each bin (Histogram). 
#### Bin size can be adjusted with the command binwidth. (1 mark)

```{r}

###############################################################################################################
# Exercise 2: Plot one variable counting the instances that fall in each bin (Histogram). 
# Bin size can be adjusted with the command binwidth. (1 mark)
###############################################################################################################

histMarksWrittenData <- sqldf("select Name, subject, Year,
                             Mark_Written
                             from studentresult
                             order by year asc")


View(histMarksWrittenData)

ggplot(histMarksWrittenData, aes(x=Mark_Written)) + 
  geom_histogram(color="red", fill="white", binwidth=25)+
  labs(y = "Count", x = "Student Marks - Written Exams") + 
  ggtitle('Histogram - Distribution of Marks in Written Exams') +
  theme(legend.position = "right",
        plot.title = element_text(size = 17),
        axis.title = element_text(size = 15),
        axis.text.x = element_text(size=15, face="bold", colour = "black"))


```







#### Exercise 3: Change the orientation of a bar chart to be horizontal rather than vertical. (1 mark)

```{r}

###############################################################################################################
# Exercise 3: Change the orientation of a bar chart to be horizontal rather than vertical. (1 mark)
###############################################################################################################


# Build SQL for all students average marks
diffWrtnOral_Score <- sqldf("select Name,
                       (sum(Mark_Written)/count(Mark_Written))-(sum(Mark_Oral)/count(Mark_Oral)) as Written_Oral_Diff
                       from studentresult
                       group by Name")

# Review aggregations of average score for students
View(diffWrtnOral_Score)


# Question 3 Bar Graphs - Horizontal Bar Chart of differences between Written and Oral Marks
gg3 <- ggplot(data=diffWrtnOral_Score, aes(x=Name, y=Written_Oral_Diff)) +
  geom_bar(stat="identity", colour="black", fill="blue") +
  labs(y = "Average % Difference", x = "Student Name") + 
  ggtitle('Average Difference Between Written and Oral Marks (per Student)') +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 14),
        axis.text.y = element_text(size=12, face="bold", colour = "black")) +
  coord_flip()
print(gg3)


```






#### Exercise 4: Part ONE: Fit a line to a scatter plot of age against Written Marks. (1 mark)

```{r}
warning=FALSE
###############################################################################################################
# Exercise 4: Part ONE: Fit a line to a scatter plot of age against Written Marks. (1 mark)
###############################################################################################################

# Convert date to a "real date" 
studentresult$DOB <- as.Date(studentresult$DOB, '%d-%m-%Y')

# Add today's date, also as a "real date" 
studentresult$today <- as.Date(Sys.Date())


# Average Result - Exercise 1
wrtnMarksByAge <- sqldf('select Name, 
                                subject, 
                                Mark_Written, 
                                floor((today-DOB)/365.25) as Age
                                from studentresult')

View(wrtnMarksByAge)


# Question 4 - Scatter Plots Pt 1
gg4a <- ggplot(data=wrtnMarksByAge, aes(x=Age, y=Mark_Written)) +
  geom_point(colour="black", size=1.5, shape=21, fill="black") +
  labs(y = "Written Exams Marks (%)", x = "Age") + 
  ggtitle('Scatter Plot of Age Against Written Marks') +
  theme(plot.title = element_text(size = 17),
        axis.title = element_text(size = 15),
        axis.text.y = element_text(size=12, face="bold", colour = "black"),
        axis.text.x = element_text(size=12, face="bold", colour = "black")) +
  geom_smooth(method='loess', se = TRUE)
print(gg4a)



```






#### Exercise 4: Part TWO: Create a scatter plot with a separate line for each data series (per subject). (1 mark)

```{r}


###############################################################################################################
# Exercise 4: Part TWO: Create a scatter plot with a separate line for each data series (per subject). (1 mark)
###############################################################################################################


# Question 4 - Scatter Plots Pt 2
gg4b <- ggplot(data=wrtnMarksByAge, aes(x=Age, y=Mark_Written, color=Subject)) +
  geom_point() +
  labs(y = "Written Exams Marks (%)", x = "Age") + 
  ggtitle('Scatter Plot of Age Against Written Marks (per Subject)') +
  theme(plot.title = element_text(size = 17),
        axis.title = element_text(size = 15),
        axis.text.y = element_text(size=12, face="bold", colour = "black"),
        axis.text.x = element_text(size=12, face="bold", colour = "black")) +
  geom_smooth(method='lm', se = FALSE, fullrange=TRUE)
print(gg4b)



```







#### Exercise 5: Create a bubble chart using ggplot by creating a scatter plot and adding the third dimension in 
#### the aesthetic option size. A fourth dimension can be added using the colour fill. (2 marks)

```{r}

###############################################################################################################
# Exercise 5: Create a bubble chart using ggplot by creating a scatter plot and adding the third dimension in 
# the aesthetic option size. A fourth dimension can be added using the colour fill. (2 marks)
###############################################################################################################

# Average Result - Exercise 1
wrtnMarksBySubject <- sqldf('select subject,
                                Year,
                                Mark_Written,
                                Mark_Oral,
                                ((Mark_Written+Mark_Oral)/2) as Subject_Mark
                                from studentresult
                                where year = 2015')

#View(wrtnMarksBySubject)



# Question 5 - Bubble Chart
gg5 <- ggplot(data=wrtnMarksBySubject, aes(x=Subject, y=Mark_Written, 
                                           size=Mark_Oral, color=Subject_Mark)) +
  geom_point(alpha=2.5) +
  scale_size(range = c(5, 13), name="Oral Results") +
  labs(y = "Written Exams Marks (%)", x = "Subject") + 
  ggtitle('Scatter Plot of Written Marks Results Against Subject (2015)') +
  theme(plot.title = element_text(size = 17),
        axis.title = element_text(size = 15),
        axis.text.y = element_text(size=12, face="bold", colour = "black"),
        axis.text.x = element_text(size=12, face="bold", colour = "black")) 
print(gg5)


```
